### 2025-02-13 Paper_Results.Rmd
update files

---
title: "Paper_Results"
output: html_document
date: "2025-01-20"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
library(dplyr)
library(lme4)
library(caret)
library(pROC)
library(patchwork)
```
## General Information
This R Markdown file can be used to replicate our results that are presented in the paper. Please note that it does not contain the shiny app.

## Pitfall 1: Within person performance

```{r }
############ Table 3 #############
##################################

# Basic Parameters from Study
n_features <- 10  # Features
n_samples <- 90  # Timepoints
n_subjects <- 150 # Subjects

# Generating Outcome
overall_prob_outcome <- 0.1  # Overall probability of 1 (for the entire dataset) (e.g., tracking depression 0.146)
sd_outcome <- 0.2 # Controls variability BETWEEN different subject (nneds to be smaller than sqrt(overall_prob_outcome*(1-overall_prob_outcome)))

# Generating Features
A <- 0.05 # Relationship between features and outcome
feature_std <- 0.1 # population level feature generating process
B <- 0.8  # Cross-subject variability ("random effect") (added per participants for all timepoints)
C <- 0.2 # Within-subject variability (added within participant for each timepoint)

source("Simulation_Functions_notshiny.R")

set.seed(12361488)
features_sample <- create_data(n_features,n_samples,n_subjects,A,feature_std,B,C,overall_prob_outcome,sd_outcome,time_effect = FALSE)
features_sample <- features_sample[[1]]

sim_1 <- run_simulation(features_sample,"row-wise",1, testsize = 0.3)
sim_2 <- run_simulation_centering(features_sample,"row-wise",1,testsize = 0.3)
sim_3 <- run_simulation(features_sample,"subject-wise",1, testsize = 0.3)
sim_4 <- run_simulation_slidingwindow(features_sample,1,windowsize = 14)

table <- rbind(sim_1$overall_summary,sim_3$overall_summary, sim_4$overall_summary)
table$'AUC Overall' <- c(sim_1$auc_value,sim_3$auc_value,sim_4$auc_value)
colnames(table) <-  c("Mean AUC within-person:", "SD AUC within-person:", "% of AUC > 0.5 within-person:", "N included within-person:", "AUC Overall")
rownames(table) <- c("row-wise","subject-wise","rolling-window")
table <- round(table[,c(5,1,2,3,4)], 2)
table
```

```{r}
############# Figure 4 #################
########################################

result <- read.csv("simulation_results.csv") # see Run_Simulation.R to investigate how results were created

p1 <- ggplot(result[result$A == 0,], aes(x=icc, y=auc, color = icc_pred)) +
  geom_point(alpha=0.4, size = 0.8) +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
  ylab("Prediction Performance (Overall AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome \n(ICC Outcome)") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  ggtitle("Bias in Performance Estimates (AUC) When No Relationship Exists: Expected AUC = 0.5") +
  geom_hline(yintercept = 0.5, color = "#83AF9B", size = 0.8) +
  annotate("text", x = 0.07, y = 0.82, label =  "Greater bias when between person differences explain a \nlarger % of total variation in both outcome and predictors.",  hjust = 0) +
  geom_curve(
    aes(x = 0.4, y = 0.85, xend = 0.68, yend = 0.87), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = -0.2
  ) + 
  annotate("text", x = 0.38, y = 0.54, label = "Low bias if between person differences explain a small \n% of total variation in outcome, predictors, or both.",  hjust = 0) +
    geom_curve(
    aes(x = 0.37, y = 0.54, xend = 0.33, yend = 0.51), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.1
  ) +
  # Adding a diagonal line to illustrate the linear relationship
  geom_abline(slope = 0.5, intercept = 0.52, color = "#FE4365",  size = 0.8)

ggsave("Figures/Example2.jpg",
  plot = p1,
  width = 10.2,
  height = 4.5)

p1

mean(result$auc_individual[result$A == 0])
sd(result$auc_individual[result$A == 0])
mean(result$auc_c[result$A == 0])
sd(result$auc_c[result$A == 0])
sd(result$auc_individual[result$A == 0])

```

```{r}
############# Figure 5 #################
########################################
result <- read.csv("simulation_results.csv")

result <- result[result$A != 0 & result$A != 0.8,]

p3 <- ggplot(result, aes(x = icc, y = auc_c_individual - auc_individual)) +
  geom_point(aes(color = ifelse(auc_c_individual - auc_individual > 0, "Above Zero", "Below or Equal to Zero")), 
             alpha = 0.3, size = 0.8, show.legend = FALSE) +
  scale_color_manual(values = c("Above Zero" = "#FC9D9A", "Below or Equal to Zero" = "#83AF9B")) +
  theme_minimal() +
  ylab("Performance Difference \n(centered - uncentered)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  ggtitle("Within-Person Performance") +
  geom_hline(yintercept = 0) + 
  annotate("text", x = 0, y = 0.28, label = "Centered predictors perform better in most cases (within-person)", hjust = 0) +
  annotate("text", x = 0, y = -0.08, label = "Uncentered predictors perform worse in most cases (within-person)", hjust = 0) +
  ylim(c(-0.45,0.35)) +
  geom_curve(
    aes(x = -0.01, y = 0.01, xend = -0.01, yend = 0.25), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#FC9D9A", 
    curvature = 0
  ) +
    geom_curve(
    aes(x = -0.01, y = 0.01, xend = -0.01, yend = -0.1), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#83AF9B", 
    curvature = 0
  )

p4 <- ggplot(result, aes(x = icc, y = auc_c - auc)) +
  geom_point(aes(color = ifelse(auc_c - auc > 0, "Above Zero", "Below or Equal to Zero")), 
             alpha = 0.3, size = 0.8,show.legend = FALSE) +
  scale_color_manual(values = c("Above Zero" = "#FC9D9A", "Below or Equal to Zero" = "#83AF9B")) +
  theme_minimal() +
  ylab("Performance Difference \n(centered - uncentered)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  ggtitle("") +
  geom_hline(yintercept = 0) + 
  annotate("text", x = 0, y = 0.28, label = "Centered predictors perform better when between person differences explain a small % of total variance (overall)", hjust = 0) +
  annotate("text", x = 0.01, y = -0.39, label = "Uncentered predictors perform better when between person differences explain a large % of total variance (overall)", hjust = 0) +
  ylim(c(-0.45,0.35)) +
   ggtitle("Overall Performance (Within and Between-Person)")  +
    geom_curve(
    aes(x = -0.01, y = 0.01, xend = -0.01, yend = 0.27), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#FC9D9A", 
    curvature = 0
  ) +
    geom_curve(
    aes(x = 0.8, y = 0.01, xend = 0.8, yend = -0.41), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.5, 
    color = "#83AF9B", 
    curvature = 0
  )

combined_plot <- (p3 / p4) +
  plot_annotation(
    title = "Performance difference: centered vs. not-centered",
  ) 
combined_plot

ggsave(  "Figures/Example4.jpg",
  plot = combined_plot,
  width = 10.2,
  height = 5.5)

```

```{r}
############# Figure 6 #################
########################################
result <- read.csv("simulation_results.csv")
result <- result[result$A != 0 & result$A != 0.8,]
p1 <- ggplot(result, aes(x=auc_c, y=auc_c_individual, color = icc_pred)) +
  geom_point(alpha=0.5, size = 1) +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
  ylab("AUC within-person centered") +
  xlab("AUC centered") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  geom_hline(yintercept = 0.5)   + 
annotate("text", x = 0.5, y = 0.9, label = "Linear relationship between overall AUC \nand within-person AUC",  hjust = 0) +
  geom_curve(
    aes(x = 0.63, y = 0.87, xend = 0.75, yend = 0.8), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = -0.15
  ) +
  ggtitle("Uncentered Predictors")

p2 <- ggplot(result, aes(x=auc, y=auc_individual, color = icc_pred)) +
  geom_point(alpha=0.5, size = 1) +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#2A363B","#83AF9B","#C8C8A9","#F9CDAD","#FC9D9A","#FE4365")) +
  ylab("AUC within-person") +
  xlab("AUC") +
  guides(col = guide_colourbar(title = "% of Total Variation Explained by \nBetween Person Differences \nin Predictors (ICC Predictor)")) +
  geom_hline(yintercept = 0.5) + 
annotate("text", x = 0.5, y = 0.9, label = "If between person differences explain a large % of variation in predictors and outcome,\nthe overall AUC becomes high while the within-person AUC remains low.",  hjust = 0) +
  geom_curve(
    aes(x = 0.75, y = 0.82, xend = 0.9, yend = 0.6), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.35
  ) +
  ggtitle("Centered Predictors")

combined_plot <- (p2 / p1) +
  plot_annotation(
    title = "Relationship Overall AUC and within-person AUC (A = 0.05 and A = 0.2)",
  ) 
combined_plot

ggsave( "Figures/Example3.jpg",
  plot = combined_plot,
  width = 11.5,
  height = 6.5)
```

## Pitfall 2: Mismatch between chosen cross-validation strategy, employment, and actual observations 
```{r }
############ Table 7 #############
##################################

# Merge the p-values with the original dataset for plotting
data <- features_sample

# Compute mean y for each subject and reorder factor levels in ascending order
subject_stats <- data %>%
  group_by(subject) %>%
  summarize(mean_y = mean(y, na.rm = TRUE),
            has_y1 = any(y == 1)) %>%  # Identify if subject has any y = 1
  arrange(mean_y)

# Convert subject to a factor with the same levels as in data
subject_stats$subject <- factor(subject_stats$subject, levels = subject_stats$subject)

# Count the number of subjects with at least one y = 1
num_subjects_with_y1 <- sum(subject_stats$has_y1)

# Apply the new factor levels to data
data <- data %>%
  mutate(subject = factor(subject, levels = subject_stats$subject))

# Merge "has_y1" information into the main dataset
data <- left_join(data, subject_stats, by = "subject")

# Create the sorted plot with conditional background color
p2 <- ggplot(data, aes(x = time, y = y)) +
  # Add background shading for subjects with y = 1
  geom_rect(
    data = data %>% filter(!has_y1),
    aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf),
    fill = "#E5E4E2", alpha = 0.05, inherit.aes = T
  ) +
  geom_point(color = "#FC9D9A", size = 0.5) +                       
  facet_wrap(~subject) +               
  labs(
    x = "Time", 
    y = "Outcome", 
    title = paste("Within-person variability over time: \n88 participants (58.6%) do not experience within-person variability. ")
  ) +
  scale_y_continuous(breaks = c(0, 1)) +  # Ensure only 0 and 1 appear on the y-axis
  theme_minimal() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  guides(fill = "none")

p2

ggsave( "Figures/Example_var.jpg",
  plot = p2,
  width = 8,
  height = 7)

```

## Pitfall 3: Weak benchmark and missing baseline comparison

```{r }
############ Table 5 #############
##################################
sim_1$auc_value_base # row-wise
sim_4$auc_value_base # rolling window

sim_4$auc_value

test <- sim_1$ind_base[[1]] %>%
  group_by(subject) %>%
  filter(length(unique(true)) > 1) %>%
  filter(sum(true == 1) > 0, sum(true == 0) > 0) %>%
  summarise(
    auc_val = auc(roc(as.numeric(as.character(true)), as.numeric(as.character(pred)), quiet = TRUE))[1],
    .groups = "drop" # Ungroup after summarizing
  )

mean(test$auc_val)

```

```{r }
############ Figure 8 #############
##################################

result = read.csv("simulation_results_baseline.csv")
p7 <- ggplot(result, aes(x=icc, y=auc_value_base)) +
  geom_point(alpha=0.5, size = 1.2, color = "#83AF9B") +
  #scale_size(range = c(.001, 6), name="") +
  theme_minimal() +
  scale_colour_gradientn(colours = c("#83AF9B")) +
  ylab("Overall Performance (AUC)") +
  xlab("% of Total Variation Explained by Between Person Differences in the Outcome (ICC Outcome)") +
  guides(col = guide_colourbar()) +
  ggtitle(paste("Performance of a Random Intercept Only Model (No Predictors Included)")) +
  geom_hline(yintercept = 0.5) +
  ylim(range= c(0.5,1)) +
    annotate("text", x = 0.5, y = 0.82, label =  "Performance increases as between person differences explain a \nlarger % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.7, y = 0.85, xend = 0.5, yend = 0.91), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.2
  ) +
      annotate("text", x = 0.2, y = 0.62, label =  "Performance can be moderate even if between person differences explain only a relatively \nsmall % of total variation in the outcome",  hjust = 0) +
  geom_curve(
    aes(x = 0.19, y = 0.62, xend = 0.07, yend = 0.61), 
    arrow = arrow(length = unit(0.08, "inch")), 
    size = 0.3, 
    color = "gray20", 
    curvature = 0.2
  )

ggsave(  "Figures/figure7.jpg",
  plot = p7,
  width = 10,
  height = 5)

```